---
title: "Traffic Stops Analysis Shared Code"
author: "Carson Groulx"
date: "4/8/2021"
output: html_document
---

Load, clean, combine data. Only necessary if using uncleaned data from Charlotte Open Data Portal. If using the provided data set, just import it and 
assign it to 'df' and skip this chunk. 
```{r}
#Load data pulled from Charlotte Open Data Portal
df1 = Officer_Traffic_Stops_Just
df2 = Officer_Traffic_Stops

#Fields of interest
fields = c("Month_of_Stop", "Reason_for_Stop", "Officer_Race", "Officer_Gender", "Officer_Years_of_Service",
           "Driver_Race", "Driver_Gender", "Driver_Age", "Was_a_Search_Conducted", "Result_of_Stop")

#All rows with missing values are removed
df1 = na.omit(df1)
df2 = na.omit(df2)

#Selecting fields of interest
df1 = df1[,fields]
df2 = df2[,fields]

#Combining data sets
df = rbind(df1, df2)

#Copy of data frame with all races included
all.races.df = df

#Restricts data frame to instances with only Black or White drivers
df = df[df$Driver_Race%in%c('Black', 'White'), ]
```

Fixes issue where values for Reason_for_Stop contain empty spaces at the end of each reason.
This issue was found in the 2019/2020 data set. If using the provided data set, skip this chunk. 
```{r}
library(stringr)

df$Reason_for_Stop = str_squish(df$Reason_for_Stop)
all.races.df$Reason_for_Stop = str_squish(all.races.df$Reason_for_Stop)
```

Common variables used throughout different chunks. Must run this before any below chunks.
```{r}
#grouping reason for stop
investigatory = c("Vehicle Regulatory", "Vehicle Equipment", "SeatBelt", "Investigation")
safety = c("Speeding", "Safe Movement", "Stop Light/Sign", "Driving While Impaired")
check = c("CheckPoint")
#indicies
i.search = df$Was_a_Search_Conducted=="Yes"
i.investigatory = df$Reason_for_Stop %in% investigatory
i.safety = df$Reason_for_Stop %in% safety
i.check = df$Reason_for_Stop %in% check
i.arrest = df$Result_of_Stop %in% c('Arrest')
i.citation = df$Result_of_Stop %in% c('Citation Issued')
i.under25 = df$Driver_Age <=25 & df$Driver_Age >=18
i.over25 = df$Driver_Age >25 & df$Driver_Age <= 65
i.black = df$Driver_Race == 'Black'
i.white = df$Driver_Race == 'White'
groups = c("Black Female", "White Female", "Black Male", "White Male")
```

Pareto Charts for driver and officer race distributions
```{r}
#install.packages("qcc")
library(qcc)

#counts of officers and drivers
n.bla.driver = sum(all.races.df$Driver_Race=='Black')
n.whi.driver = sum(all.races.df$Driver_Race=='White')
n.asi.driver = sum(all.races.df$Driver_Race=='Asian')
n.nat.driver = sum(all.races.df$Driver_Race=='Native American')

n.2or.officer = sum(all.races.df$Officer_Race=='2 or More')
n.ame.officer = sum(all.races.df$Officer_Race=='American Indian/Alaska Native')
n.asi.officer = sum(all.races.df$Officer_Race=='Asian / Pacific Islander')
n.his.officer = sum(all.races.df$Officer_Race=='Hispanic/Latino')
n.nat.officer = sum(all.races.df$Officer_Race=='Native Hawaiian/Oth Pac Island')
n.not.officer = sum(all.races.df$Officer_Race=='Not Specified')
n.bla.officer = sum(all.races.df$Officer_Race=='Black/African American')
n.whi.officer = sum(all.races.df$Officer_Race=='White')

driver.races = c("Black", "White", "Asian", "Native American")
officer.races = c('2+', 'Ame.Ind./Ala.N.', 'Asian/Pac. Islander', 'Black', 'Hispanic/Latino', 'N.Haw./Other Pac. Isl.', 'Not Specified', 'White')

driver.counts = c(n.bla.driver, n.whi.driver, n.asi.driver, n.nat.driver)
officer.counts = c(n.2or.officer, n.ame.officer, n.asi.officer, n.bla.officer, n.his.officer, n.nat.officer, n.not.officer, n.whi.officer)

driver.pct = round(driver.counts/sum(driver.counts)*100)
officer.pct = round(officer.counts/sum(officer.counts)*100)

driver.races = paste(driver.races, driver.pct) 
driver.races = paste(driver.races,"%")
officer.races = paste(officer.races, officer.pct) 
officer.races = paste(officer.races,"%") 

names(driver.counts) = driver.races
names(officer.counts) = officer.races

pareto.chart(driver.counts, main = "Driver proportions by race", col = c(1:5))
pareto.chart(officer.counts, main = "Officer proportions by race", col = c(1:8))
```

Reason for stop
```{r}
races = c("Black", "White")
colors = c(1:length(races))

#Bar chart for proportions of stops with a search conducted. Bases are within each racial group.
reasons.base = table(df$Driver_Race[i.investigatory |
i.safety | i.check])
reasons.investigatory = table(df$Driver_Race[i.investigatory])
reasons.safety = table(df$Driver_Race[i.safety])
reasons.check = table(df$Driver_Race[i.check])

investigatory.p = reasons.investigatory / reasons.base
safety.p = reasons.safety / reasons.base
check.p = reasons.check / reasons.base

mat.reasons.p = matrix(c(investigatory.p, safety.p, check.p), ncol = 3)

barplot(mat.reasons.p, beside = T, col = colors,
        names.arg = c("Investigatory", "Safety", "Checkpoint"),
        ylab = "Proportions of stops", main = "Reason for stop",
        legend = races, 
        args.legend = list(title = "Races", x = "topright", pch = 15, col = colors, bty = 'n')
)

#Bar chart for proportions above compared: Black/White
investigatory.compared = investigatory.p['Black'] / investigatory.p['White']
safety.compared = safety.p['Black'] / safety.p['White']
check.compared = check.p['Black'] / check.p['White']

mat.reasons.compare = matrix(c(investigatory.compared, safety.compared, check.compared), ncol = 3)

barplot(mat.reasons.compare, beside = T, col = 'black',
        names.arg = c("Investigatory", "Safety", "Checkpoint"),
        ylab = "Proportions", xlab = "Black / White", main = "Reasons for Stops Compared")

abline(h = c(1))

# hyp testing for above
reasons.investigatory.black = reasons.investigatory["Black"]
reasons.investigatory.white = reasons.investigatory["White"]
reasons.base.black = reasons.base['Black']
reasons.base.white = reasons.base['White']

print('Investigatory')
prop.test(
x = c(reasons.investigatory.black, reasons.investigatory.white),
n = c(reasons.base.black, reasons.base.white)
)

reasons.safety.black = reasons.safety["Black"]
reasons.safety.white = reasons.safety["White"]

print('Safety')
prop.test(
x = c(reasons.safety.black, reasons.safety.white),
n = c(reasons.base.black, reasons.base.white)
)

reasons.check.black = reasons.check["Black"]
reasons.check.white = reasons.check["White"]

print('Checkpoint')
prop.test(
x = c(reasons.check.black, reasons.check.white),
n = c(reasons.base.black, reasons.base.white)
)

#Broken down by gender
colors = c('yellow', 'purple', 'green', 'red')

reasons.base.genders = table(df$Driver_Race[i.investigatory | i.safety | i.check], 
                             df$Driver_Gender[i.investigatory | i.safety | i.check])

investigatory.genders.p = table(df$Driver_Race[i.investigatory], df$Driver_Gender[i.investigatory]) / reasons.base.genders
safety.genders.p = table(df$Driver_Race[i.safety], df$Driver_Gender[i.safety]) / reasons.base.genders
check.genders.p = table(df$Driver_Race[i.check], df$Driver_Gender[i.check]) / reasons.base.genders

mat.reasons.genders.p = matrix(c(investigatory.genders.p, safety.genders.p, check.genders.p), ncol = 3)

barplot(mat.reasons.genders.p, beside = T, col = colors,
        names.arg = c("Investigatory", "Safety", "Checkpoint"),
        ylab = "Proportions of stops", main = "Reason for stop",
        legend = groups,
        args.legend = list(x = "topright", pch = 15, col = colors, bty = 'n')
)

#Above compared: other/Black males
investigatory.genders.compared = c(
investigatory.genders.p['Black', 'Female'] / investigatory.genders.p['Black', 'Male'],
investigatory.genders.p['White', 'Female'] / investigatory.genders.p['Black', 'Male'],
investigatory.genders.p['Black', 'Male'] / investigatory.genders.p['Black', 'Male'],
investigatory.genders.p['White', 'Male'] / investigatory.genders.p['Black', 'Male']
)

safety.genders.compared = c(
safety.genders.p['Black', 'Female'] / safety.genders.p['Black', 'Male'],
safety.genders.p['White', 'Female'] / safety.genders.p['Black', 'Male'],
safety.genders.p['Black', 'Male'] / safety.genders.p['Black', 'Male'],
safety.genders.p['White', 'Male'] / safety.genders.p['Black', 'Male']
)

check.genders.compared = c(
check.genders.p['Black', 'Female'] / check.genders.p['Black', 'Male'],
check.genders.p['White', 'Female'] / check.genders.p['Black', 'Male'],
check.genders.p['Black', 'Male'] / check.genders.p['Black', 'Male'],
check.genders.p['White', 'Male'] / check.genders.p['Black', 'Male']
)

mat.reasons.genders.compare = matrix(c(investigatory.genders.compared, 
                                       safety.genders.compared, 
                                       check.genders.compared), 
                                     ncol = 3)

barplot(mat.reasons.genders.compare, beside = T, col = colors,
        names.arg = c("Investigatory", "Safety", "Checkpoint"),
        ylab = "Proportions", xlab = "Other / Black Male", main = "Reason for stop Compared",
        legend = groups,
        args.legend = list(x = "topleft", pch = 15, col = colors, bty = 'n', cex = .8)
)

#reduced to stops with searches
#omitted checkpoints due to infrequency
reasons.base.searches = table(df$Driver_Race[(i.investigatory | i.safety) & i.search], 
                              df$Driver_Gender[(i.investigatory | i.safety) & i.search])

investigatory.searches.p = table(df$Driver_Race[i.investigatory & i.search], 
                                 df$Driver_Gender[i.investigatory & i.search]) / reasons.base.searches

safety.searches.p = table(df$Driver_Race[i.safety & i.search], 
                          df$Driver_Gender[i.safety & i.search]) / reasons.base.searches

mat.reasons.searches.p = matrix(c(investigatory.searches.p, safety.searches.p), ncol = 2)

barplot(mat.reasons.searches.p, beside = T, col = colors,
        names.arg = c("Investigatory", "Safety"),
        ylab = "Proportions of stops", main = "Reason for stop in searches",
        legend = groups,
        args.legend = list(x = "topright", pch = 15, col = colors, bty = 'n')
)

#above compared: other/ Black Male
investigatory.searches.genders.compared = c(
investigatory.searches.p['Black', 'Female'] / investigatory.searches.p['Black', 'Male'],
investigatory.searches.p['White', 'Female'] / investigatory.searches.p['Black', 'Male'],
investigatory.searches.p['Black', 'Male'] / investigatory.searches.p['Black', 'Male'],
investigatory.searches.p['White', 'Male'] / investigatory.searches.p['Black', 'Male']
)

safety.searches.genders.compared = c(
safety.searches.p['Black', 'Female'] / safety.searches.p['Black', 'Male'],
safety.searches.p['White', 'Female'] / safety.searches.p['Black', 'Male'],
safety.searches.p['Black', 'Male'] / safety.searches.p['Black', 'Male'],
safety.searches.p['White', 'Male'] / safety.searches.p['Black', 'Male']
)

mat.searches.genders.compared = matrix(c(investigatory.searches.genders.compared, 
                                         safety.searches.genders.compared), 
                                       ncol = 2)

barplot(mat.searches.genders.compared, beside = T, col = colors,
        names.arg = c("Investigatory", "Safety"),
        ylab = "Proportions of stops", xlab = "Other / Black Male", main = "Reason for stop in searches Compared",
        legend = groups,
        args.legend = list(x = "topleft", pch = 15, col = colors, bty = 'n')
)
```

Results of Stops
```{r}
results = c('Arrest', 'Citation Issued', 'No Action Taken',  'Verbal Warning', 'Written Warning')
colors = c(1:length(results))

#Bar charts for results of stop (counts), grouped by race only
results.black = table(df$Result_of_Stop[i.black])
results.white = table(df$Result_of_Stop[i.white])

mat.results.race = matrix(c(results.black, results.white), ncol = 2)

barplot(mat.results.race, beside = T, col = colors,
        names.arg =  c("Black", "White"),
        xlab = "Result of stop",
        ylab = "Count of stops", 
        main = "Result of stop (counts)",
        legend = results,
        args.legend = list(title = "Results", x = "topleft", pch = 15, col = colors, bty = 'n', cex = .60))

#Bar charts for results of stop (proportions), grouped by race only
results.black.p = results.black / sum(results.black)
results.white.p = results.white / sum(results.white)

mat.results.race.p = matrix(c(results.black.p, results.white.p), ncol = 2)

barplot(mat.results.race.p, beside = T, col = colors,
        names.arg =  c("Black", "White"),
        xlab = "Result of stop",
        ylab = "Proportions", 
        main = "Result of stop (Proportions)",
        legend = results,
        args.legend = list(title = "Results", x = "topleft", pch = 15, col = colors, bty = 'n', cex = .60))

#Bar charts for results of stop compared White / Black (proportions), grouped by race only

results.race.compared = results.white.p / results.black.p

barplot(results.race.compared, beside = T, col = colors,
        names.arg =  c("White / Black"),
        xlab = "Result of stop",
        ylab = "Proportions", 
        main = "Result of stop Compared (Proportions)",
        legend = results,
        args.legend = list(title = "Results", x = "topleft", pch = 15, col = colors, bty = 'n', cex = .60))
abline(h=c(1))

#Bar chart of total counts of each result of stop, grouped by race and gender
results.black.gender = table(df$Result_of_Stop[i.black], df$Driver_Gender[i.black])
results.white.gender = table(df$Result_of_Stop[i.white], df$Driver_Gender[i.white])

results.black.female = results.black.gender[,"Female"]
results.white.female = results.white.gender[,"Female"]
results.black.male = results.black.gender[,"Male"]
results.white.male = results.white.gender[,"Male"]

mat.results.race.gender = matrix(c(results.black.female, results.white.female, results.black.male, results.white.male), ncol = 4)

barplot(mat.results.race.gender, beside = T, col = colors,
        names.arg =  groups,
        xlab = "Result of stop",
        ylab = "Count of stops", 
        main = "Result of stop (counts)",
        legend = results,
        args.legend = list(title = "Results", x = "topleft", pch = 15, col = colors, bty = 'n', cex = .60))


#Bar chart of proportions of each result of stop, grouped by race and gender
results.black.female.p = results.black.gender[,'Female']/sum(results.black.gender[,'Female'])
results.white.female.p = results.white.gender[,'Female']/sum(results.white.gender[,'Female'])
results.black.male.p = results.black.gender[,'Male']/sum(results.black.gender[,'Male'])
results.white.male.p = results.white.gender[,'Male']/sum(results.white.gender[,'Male'])

mat.results.race.gender.p = matrix(c(results.black.female.p, results.white.female.p,results.black.male.p, results.white.male.p), ncol = 4)

barplot(mat.results.race.gender.p, beside = T, col = colors,
        names.arg = groups,
        xlab = "Result of stop",
        ylab = "Proportions", 
        main = "Results of stop (proportions)",
        legend = results,
        args.legend = list(title = "Results", x = "topleft", pch = 15, col = colors, bty = 'n', cex = .60))

#above compared: other / Black male
mat.results.race.gender.compared = matrix(c(results.black.female.p / results.black.male.p,
                                results.white.female.p / results.black.male.p,
                                results.black.male.p / results.black.male.p,
                                results.white.male.p / results.black.male.p), ncol = 4)

barplot(mat.results.race.gender.compared, beside = T, col = colors,
        names.arg = groups,
        xlab = "Other / Black male",
        ylab = "Proportions", 
        main = "Results of stop compared",
        legend = results,
        args.legend = list(title = "Results", x = "topleft", pch = 15, col = colors, bty = 'n', cex = .60))
abline(h=c(1))

```

Searches
```{r}
races = c("Black", "White")
colors = c(1:length(races))

#Bar chart for stops with a search conducted (counts)
base.race = table(df$Driver_Race)

search.race = table(df$Driver_Race[i.search])
search.race.investigatory = table(df$Driver_Race[i.search & i.investigatory])
search.race.safety = table(df$Driver_Race[i.search & i.safety])
search.race.check = table(df$Driver_Race[i.search & i.check])

mat.search.race.reason = matrix(c(search.race, search.race.investigatory, 
                                    search.race.safety, search.race.check), 
                                  ncol = 4)

barplot(mat.search.race.reason, beside = T, col = colors,
        names.arg = c("All", "Investigatory", "Safety", "Checkpoint"),
        xlab = "Reason for stop",
        ylab = "Proportions",
        main = "Search rates (Counts)",
        legend = races,
        args.legend = list(title = "Races", x = "topright", pch = 15, col = colors, bty = 'n'))

#Bar chart for stops with a search conducted (proportions) 
search.race.p = search.race/base.race
search.race.investigatory.p = search.race.investigatory/base.race
search.race.safety.p = search.race.safety/base.race
search.race.check.p = search.race.check/base.race

mat.search.race.reason.p = matrix(c(search.race.p, search.race.investigatory.p, 
                                    search.race.safety.p, search.race.check.p), 
                                  ncol = 4)

barplot(mat.search.race.reason.p, beside = T, col = colors,
        names.arg = c("All", "Investigatory", "Safety", "Checkpoint"),
        xlab = "Reason for stop",
        ylab = "Proportions",
        main = "Search rates (Proportions)",
        legend = races,
        args.legend = list(title = "Races", x = "topright", pch = 15, col = colors, bty = 'n'))

#Bar chart for proportions above compared. White / Black
races = c("Black/White")
colors = c(1:length(races))

search.race.compare = search.race.p['White']/search.race.p['Black']
search.race.investigatory.compare = search.race.investigatory.p['White']/search.race.investigatory.p['Black']
search.race.safety.compare = search.race.safety.p['White']/search.race.safety.p['Black']
search.race.check.compare = search.race.check.p['White']/search.race.check.p['Black']
mat.search.race.reason.compare = matrix(c(search.race.compare, search.race.investigatory.compare,
                                          search.race.safety.compare, search.race.check.compare), ncol = 4)

barplot(mat.search.race.reason.compare, beside = T, col = colors,
        names.arg = c("All", "Investigatory", "Safety", "Checkpoint"),
        xlab = "White / Black",
        ylab = "Proportions",
        main = "Search Rates Compared")


#Bar charts for proportions of searches with Arrest
races = c("Black", "White")
colors = c(1:length(races))

search.race.arrest.p = (table(df$Driver_Race[i.search & i.arrest]))/
          (table(df$Driver_Race[i.search]))

search.race.arrest.investigatory.p = (table(df$Driver_Race[i.search & i.arrest & i.investigatory]))/
          (table(df$Driver_Race[i.search & i.investigatory]))

search.race.arrest.safety.p = (table(df$Driver_Race[i.search & i.arrest & i.safety]))/
          (table(df$Driver_Race[i.search & i.safety]))

search.race.arrest.check.p = (table(df$Driver_Race[i.search & i.arrest & i.check]))/
          (table(df$Driver_Race[i.search & i.check]))

mat.search.arrest.race.reason.p = matrix(c(search.race.arrest.p, search.race.arrest.investigatory.p, 
                    search.race.arrest.safety.p, search.race.arrest.check.p), ncol = 4)

barplot(mat.search.arrest.race.reason.p, beside = T, col = colors,
        names.arg = c("All", "Investigatory", "Safety", "Checkpoint"),
        xlab = "Reason for stop",
        ylab = "Proportions", ylim = c(0, 1),
        main = "Arrest Rate with Search",
        legend = races,
        args.legend = list(title = "Races", x = "topleft", pch = 15, col = colors, bty = 'n', cex = .8))

#Bar charts for proportions of searches with Citations
races = c("Black", "White")
colors = c(1:length(races))

search.race.citation.p = (table(df$Driver_Race[i.search & i.citation]))/
          (table(df$Driver_Race[i.search]))

search.race.citation.investigatory.p = (table(df$Driver_Race[i.search & i.citation & i.investigatory]))/
          (table(df$Driver_Race[i.search & i.investigatory]))

search.race.citation.safety.p = (table(df$Driver_Race[i.search & i.citation & i.safety]))/
          (table(df$Driver_Race[i.search & i.safety]))

search.race.citation.check.p = (table(df$Driver_Race[i.search & i.citation & i.check]))/
          (table(df$Driver_Race[i.search & i.check]))

mat.search.citation.race.reason.p = matrix(c(search.race.citation.p, search.race.citation.investigatory.p, 
                    search.race.citation.safety.p, search.race.citation.check.p), ncol = 4)

barplot(mat.search.citation.race.reason.p , beside = T, col = colors,
        names.arg = c("All", "Investigatory", "Safety", "Checkpoint"),
        xlab = "Reason for stop",
        ylab = "Proportions", ylim = c(0, 1),
        main = "Citation Rate with Search",
        legend = races,
        args.legend = list(title = "Races", x = "topleft", pch = 15, col = colors, bty = 'n', cex = .8))



#Bar charts for proportions of searches with Arrest. Include gender and age
colors = c('yellow','green','purple','red')

#all
base.race.gender = table(df$Driver_Race, df$Driver_Gender)
search.race.gender = table(df$Driver_Race[i.search], df$Driver_Gender[i.search])

search.race.gender.p = search.race.gender/base.race.gender
arrest.race.gender.p = (table(df$Driver_Race[i.arrest], df$Driver_Gender[i.arrest])/search.race.gender)

search.arrest.race.gender.p = (table(df$Driver_Race[i.search & i.arrest], df$Driver_Gender[i.search & i.arrest])/search.race.gender)

search.citation.race.gender.p = (table(df$Driver_Race[i.search & i.citation], df$Driver_Gender[i.search & i.citation])/search.race.gender)

#18-25
search.race.gender.under25 = table(df$Driver_Race[i.search & i.under25], df$Driver_Gender[i.search & i.under25])
search.race.gender.under25.p = search.race.gender.under25/table(df$Driver_Race[i.under25], df$Driver_Gender[i.under25])

search.arrest.race.gender.under25.p = (table(df$Driver_Race[i.search&i.arrest&i.under25], df$Driver_Gender[i.search&i.arrest&i.under25]))/search.race.gender.under25

search.citation.race.gender.under25.p = (table(df$Driver_Race[i.search&i.citation&i.under25], df$Driver_Gender[i.search&i.citation&i.under25]))/search.race.gender.under25
#26-65
search.race.gender.over25 = table(df$Driver_Race[i.search & i.over25], df$Driver_Gender[i.search & i.over25])
search.race.gender.over25.p = search.race.gender.over25/table(df$Driver_Race[i.over25], df$Driver_Gender[i.over25])

search.arrest.race.gender.over25.p = (table(df$Driver_Race[i.search&i.arrest&i.over25], df$Driver_Gender[i.search&i.arrest&i.over25]))/search.race.gender.over25

search.citation.race.gender.over25.p = (table(df$Driver_Race[i.search&i.citation&i.over25], df$Driver_Gender[i.search&i.citation&i.over25]))/search.race.gender.over25

#searches by race gender age
mat.search.race.gender.age.p = matrix(c(search.race.gender.p, search.race.gender.under25.p, search.race.gender.over25.p), ncol = 3)

barplot(mat.search.race.gender.age.p, beside = T, main = 'Search Rates', col = 1:4, names.arg = c('All', '18-25', '26-65'), legend = groups,
       args.legend = list(title = "Groups", x = "topright", pch = 15, col = 1:4, bty = 'n'))

#searches with arrest by race gender age
mat.search.arrest.race.gender.age.p = matrix(c(search.arrest.race.gender.p, search.arrest.race.gender.under25.p, search.arrest.race.gender.over25.p), ncol = 3)

barplot(mat.search.arrest.race.gender.age.p, beside = T, main = 'Arrest Rates in Searches',  col = 1:4, names.arg = c('All', '18-25', '26-65'), 
legend = groups,       args.legend = list(title = "Genders", x = 'top', pch = 15, col = 1:4, bty = 'n'))

#searches with citation by race gender age
mat.search.citation.race.gender.age.p = matrix(c(search.citation.race.gender.p, search.citation.race.gender.under25.p, search.citation.race.gender.over25.p), ncol = 3)

barplot(mat.search.citation.race.gender.age.p, beside = T,  col = 1:4, main = 'Citation Rates in Searches', names.arg = c('All', '18-25', '26-65'), 
        legend = groups, args.legend = list(title = "Genders", x = "topright", pch = 15, col = 1:4, bty = 'n'))

```

Proportions of each driver race stopped binned on officer's years of service. Here I choose to include only up to year 31 because years after 31 contain so few instances.
```{r}
#Values subject to change depending on how many years of service will be charted and which races are to be charted 
startYear = 1
endYear = 31
races = c("Black","White")

#Code from here below should not need change, unless to manipulate the charts or to produce more or less charts in the instance that the races change
years = str_split(startYear:endYear, ",")
servicebyrace = matrix(nrow = endYear, ncol = length(races))
colnames(servicebyrace) = races

for (i in(1:endYear)) {
  for(j in(1:length(races))){
    servicebyrace[i, j] = sum(df$Driver_Race == races[j] & df$Officer_Years_of_Service == years[i])/sum(df$Officer_Years_of_Service == years[i])
  }
}

#plot for black and white drivers
plot(c(startYear:endYear), servicebyrace[startYear:endYear, "Black"], 
     xlim = c(startYear, endYear), ylim = c(.2,.8), 
     xlab = 'Years of Service', ylab = 'Proportion of stops', main = "Fig 4: Driver's race binned on officer years of service", 
     col = 'blue')
lines(c(startYear:endYear), servicebyrace[startYear:endYear, "Black"], col = 'blue')
abline(lm(servicebyrace[startYear:endYear, "Black"]~c(startYear:endYear)), col = 'blue', lty = 4)
points(c(startYear:endYear), servicebyrace[startYear:endYear, "White"], col = 'green')
lines(c(startYear:endYear), servicebyrace[startYear:endYear, "White"], col = 'green')
abline(lm(servicebyrace[startYear:endYear, "White"]~c(startYear:endYear)), col = 'green', lty = 4)
legend("topright", c("Black","White"), col = c('blue', 'green'), pch = 15, bty = 'n')

```

Logistic Regression
```{r}
library("ROCR")
library('caTools')
library('pscl')
library('dominanceanalysis')
library('blorr')
library('Metrics')

#Recoding and restricting data set.
#Officers are White or non-White
#reasons for stop are only investigatory or safety. Others are removed.
ndf = df[df$Reason_for_Stop%in%investigatory|df$Reason_for_Stop%in%safety, ]

i.ndf.investigatory = ndf$Reason_for_Stop%in%investigatory
i.ndf.safety = ndf$Reason_for_Stop%in%safety

ndf$Reason_for_Stop[i.ndf.investigatory] = 'Investigatory'
ndf$Reason_for_Stop[i.ndf.safety] = 'Safety'

ndf$Officer_Race[ndf$Officer_Race!='White'] = "Not White"

#Variables converted to factor types
ndf$Was_a_Search_Conducted = as.factor(ndf$Was_a_Search_Conducted)
ndf$Driver_Race = as.factor(ndf$Driver_Race)
ndf$Driver_Gender = as.factor(ndf$Driver_Gender)
ndf$Result_of_Stop = as.factor(ndf$Result_of_Stop)
ndf$Reason_for_Stop = as.factor(ndf$Reason_for_Stop)
ndf$Officer_Race = as.factor(ndf$Officer_Race)
ndf$Officer_Gender = as.factor(ndf$Officer_Gender)

#predicting search conducted
lm = glm(Was_a_Search_Conducted~relevel(Driver_Race, 'White')+Driver_Gender+Reason_for_Stop+Officer_Race+Officer_Gender+Officer_Years_of_Service+Driver_Age, data = ndf, family = binomial(link='logit'))
exp(coef(lm))
summary(lm)
anova(lm, test = 'Chisq')

#model accuracy
split = sample.split(ndf$Was_a_Search_Conducted, SplitRatio = .8)
train = ndf[split,]
test = ndf[!split, ]

lm.train = glm(Was_a_Search_Conducted~Driver_Race+Driver_Gender+Reason_for_Stop+Officer_Race+Officer_Gender+Officer_Years_of_Service+Driver_Age, data = train, family = binomial(link='logit'))

fit.res = predict(lm.train, newdata = test, type = 'response')
pr = prediction(fit.res, test$Was_a_Search_Conducted)
prf = performance(pr, measure = 'tpr', x.measure = 'fpr')
plot(prf)
auc = performance(pr, measure = 'auc')
auc = auc@y.values[[1]]
auc
```




